\documentclass{article}

\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools,thmtools,cleveref}
\usepackage{stmaryrd} % double bracket [[1, n]]
\usepackage{graphicx,dsfont,float,booktabs,multirow}

\newcommand\halfopen[2]{\ensuremath{[#1,#2)}}
\newcommand\openhalf[2]{\ensuremath{(#1,#2]}}

\begin{document}

\section{Online autonomous intersection control as
  an adaptive trajectory optimization problem}

We consider a transportation system with only autonomous vehicles that are all
connected to a single, central controller, that is in charge with steering each
individual vehicle.
%
Assume that routes of vehicles are fixed, so that vehicles can be modeled as
double integrators.
%
When no more additional vehicles arrive to the system, the coordination problem
can be stated as a trajectory optimization problem.

We want to extend this model such that it captures the random arrival of
vehicles to the network. More generally, we will allow arrivals to be announced
before the actual vehicle enters the sytem.

\paragraph{Online model.}
This setting is naturally captured as a semi-Markov decision process (SMDP).
Since we are dealing with vehicle trajectories, this means that we have
infinite-dimensional state and action spaces.
%
Let $\{ X_m, m \geq 0 \}$ denote the state process, let $\Upsilon_m$ denote the
sojourn time.

\paragraph{States.}
The state is defined as the tuple $X_m = (\mathcal{N}_m, S_m)$, where
$\mathcal{N}_m$ denotes the set of vehicle indices of the vehicles currently in
the system, and $S_m : \mathcal{N}_m \rightarrow \mathbb{R}^2$ maps each vehicle
index $i$ to the state $S_m(i) = (p_i, v_i)$ of that vehicle, consisting of its
position $p_i$ along the route and its current speed $v_i$.

\paragraph{Actions.}
Each action prescribes the trajectories for the vehicles $\mathcal{N}_m$ to
follow through the network.
%
Each action is a function
$a_m : \mathcal{N}_m \rightarrow (\halfopen{t_m}{\infty} \rightarrow \mathbb{R}^2)$

\paragraph{Constraints.}

\begin{enumerate}
  \item effect of collision avoidance on arrivals
  \item collision avoidance restricts actions
  \item continuity requirement of actions: trajectories must be a smooth
        continuation of the current trajectory plan (second-order or higher
        continuity)
\end{enumerate}

\subsection{Problem aspects}

\paragraph{Optimal control.}

\paragraph{Online scheduling.}

\subsection{Solution methods}



\end{document}
